Liveness-Based Garbage Collection for Lazy Languages

### 2015-07-11 paper/paper.tex
proof.

\documentclass[9pt]{sigplanconf}
\usepackage{etex}
\usepackage{float}
\usepackage{mathrsfs}
\usepackage{booktabs}
\usepackage{boxedminipage}
\usepackage[T1]{fontenc}
\usepackage{epsfig}
\usepackage{multirow}
\usepackage{url}
\usepackage[normalem]{ulem}
\usepackage{schemepgm}
\usepackage{graphicx}
\usepackage{graphpap}
\usepackage{tabularx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{pstricks}
\usepackage{pst-text}
\usepackage{pst-node}
\usepackage{pst-tree}
\usepackage{pst-rel-points}
\usepackage{bcprules}
\usepackage{subfigure}
\usepackage[boxed]{algorithm2e}
\usepackage{makecell}

@@

\mycomment{\color{red}
\subsection{Correctness of liveness analysis}

We shall now  outline a proof of correctness  of the liveness analysis
presented  in  Section~\ref{sec:liveness-analysis}.   This requires  a
significant modification  to the technique  used in \cite{asati14lgc}.
We shall first modify the standard semantics. This semantics, which we
shall  call {\em  minefield semantics},  will model  a  liveness based
garbage collection at  each transition. This is done  by starting from
the root-set  and inserting a  special value $\bot$ at  each reference
location in the  heap that is reachable but not  live by our analysis.
Attempts  to dereference  such  locations during  the transition  will
result in entering a special  state denoted \bang.  We shall then show
that no program enters the \bang\ state.

To set up the minefield semantics, we follow these steps:
\begin{enumerate}
\item  Given an  demand transformer  \Lfonly\ we  enrich  the abstract
  machine state $\rho, S, H, e$  to $\rho, S, H, e, \sigma$.  $\sigma$
  is the  demand on the  currently active function. As  a consequence,
  the liveness $\Lv$ at $e$ is $\mathcal{L}(e, \sigma, \Lfonly)$. Each
  entry  in   the  stack  $S$   of  suspended  evaluations,   is  also
  augmented---it is  now $(\rho, x,  e, eps, \sigma)$. The $\sigma$  in the
  stack entry  represents the demand  on the  function where  the evaluation
  \ldots.  And $eps$ is \ldots.
\item We introduce a function $GC$ that takes the 4-tuple $(\rho, H, S,
  \sigma)$ and  inserts $\bot$ in all locations in the current
  environment $\rho$, the heap $H$ and the environments in the stack
  $S$ that are dead. More formally:

and $\Sigma$ is a  stack of demands---one for each function
  frame pushed in $S$.  Second,  we augment $Val$ with a value $\bot$.
  To model strong liveness $\bot$ may be copied freely, including into
  a  \CONS\  cell,  but  explodes  when  used  computationally  (in  a
  primitive operation  other than a copy). Additionally  we define $GC
  (\Lv, \Sigma)  : (\rho,  H, S) \rightarrow  (\rho ,  H , S  )$ which
  determines live-reachability using $\rho$ and the $\rho$'s in $S$ as
  the root  set and following links in  $H$ only as far  as allowed by
  \Lv\   and   $\Sigma$.     Hence   $GC   (\Lv,   \Sigma)$   replaces
  live-unreachable  values---in $\rho$,  in  $H$ and  the $\rho$'s  in
  $S$---with $\bot$.  For example,  if $x.\epsilon \in \Lv$ then $\rho
  (x) = \bot$.  Only $GC (\cdots )$ introduces $\bot$.
\end{enumerate}

 }

@@

%Add graphs for memory usage LGC vs RGC
%Add graphs/table showing the GC times and other statistics

%Mention lower peak memory usage.
%Mention faster GC than full LGC and more garbage collected than RGC.
%Pitch it as sweet spot between full LGC and RGC.
%Future work
%Mention full LGC
%Mention k-liveness, use length example to show 1-liveness is not
%sufficient.
%Mention mixed mode GC,
% - doing RGC most of the times and switching to LGC in case RGC fails
%to collect any garbage
% - doing LGC most of the times but regularly checking the reachability
%of the heap and
% switching to RGC if LGC almost equals RGC

\bibliography{fun_hra}{}
\bibliographystyle{abbrv}
\end{document}

\pagebreak
\section{Appendix}

To set up the minefield semantics, we follow these steps:
\begin{enumerate}
\item We enrich  the abstract machine state $\rho,$ $S,$  $ H,$ $e$ to
  $\rho,$ $S,$ $H,$ $e,$ $\sigma,$ $e_f$, where $\sigma$ is the demand
  on the expression $e$, and $e_f$  is the body of the function. $e_f$
  and $\sigma$ are  both needed to calculate  the liveness environment
  for expressions.  The information  in a suspended evaluation context
  is also similarly  augmented with the demand on it  and the function
  body of  which the suspended evaluation  is a part.  Thus  a stacked
  entry now takes the form $(\rho, x, e, \sigma, e_f)$. A modification
  of the small  step semantics which carries the  extra information is
  shown in Figure~\ref{fig:minefield-semantics}.
\item Given  the current  evaluation context $\rho,  S, H,  e, \sigma,
  e_f$ and $LF$, we can construct a liveness environment as follows:
  \begin{enumerate}
  \item If $e$ is an expression, then:  
  $$ \Lv = \bigcup_{\psi \in \mathcal{P}(e)} \mathcal{M}(\psi),
    \text{where }
  \mathcal{M} = \mathcal{L}(e_f, \sigma, \Lfonly)$$
\item If $e$ is an application, then:  
  $$ \Lv = ref(e, \sigma, \Lfonly)$$
  \end{enumerate}
We can define in the same way the liveness environment for each of the
suspended  evaluation contexts in  $S$  giving  a  stack of  liveness
environments that we shall denote $\mathsf{SL}$.
\item $GC$ models a liveness based garbage collection:
$GC(\rho,$ $ S,$ $ H,$ $ L,$ $ SL) = (\rho', S', H')$ 
  \begin{enumerate}
  \item For each $x \in domain(\rho)$,  $\rho'(x)=\bot$ iff $x.\epsilon
    \notin \Lv$.
\item For each stack entry $(\rho,\_,\_,\_,\_)$ in $S$ with $\Lv'$ as
  the corresponding liveness environment in $\mathsf{SL}$,  and for each $x
  \in domain(\rho)$,  $\rho'(x)=\bot$
  iff $x.\epsilon \notin \Lv'$.
\item For each location $\ell$, $H'(\ell) = \bot$, iff  there is no
  $x$ in either the current environment or one of the stacked
  environments $\rho$ such that for some $\alpha$, $H[x.\alpha] =
  \ell$  and $x.\alpha \in \Lv$, .
  \end{enumerate}
\end{enumerate}

Starting   from  the   initial  state   $([\;]_\rho,([\;]_\rho,  \ans,
(\print~\ans)):[\;]_{S} ,  [\;]_{H}, (\mainpgm)$, every  transition is
preceded by  a garbage  collection using $GC$.  We consider  a garbage
collection followed  by a transition as  a step. We show  by induction
that, starting  from the  initial state, any  transition of  $n$ steps
occurs without a \bang.

We observe that  the transitions involving dereferencing  and can lead
to \bang are those which have  a $H(...)$ term in the transition rule.
Of these most  of the terms are of the  form $H(\rho(x))$, involving a
root variable.  There is exactly  one occurrence of $H(l)$ which means
dereferencing using a heap location.

In both  the cases  we show  that the dereferencing  cannot lead  to a
\bang\ state in a function call less program.

For this we need the following lemma,
\begin{lemma}
\label{lem:correctness}
Consider, 
\begin{enumerate}
\item
 A \LET\ definition $\pi_i:(\LET~x \leftarrow~s~\IN~e)$ such
that there is  an evaluation path in $e$ in  which $\pi_i$ is followed
by a  use i.e an  occurrence of $\pi_j:(\CAR\ x)$,  $\pi_j:(\CDR\ x)$,
$\pi_j:(+\ x\ y)$, $\pi_j:(\SIF\ x\ e1\ e2)$ or $\pi_j:(\SRETURN\ x)$
\item
 A \LET\ definition $\pi_i(\LET~z  \leftarrow~(\CONS~x~y)~\IN~e$, such that
 $e$ includes  an evaluation path in  which there is an  occurrence of
 $\pi_j: (\CAR~z)$ and $\pi_j$ follows $\pi_i$.
\end{enumerate},
the demand on $x$ is non-null between $\pi_i$ and $\pi_j$.
\end{lemma} 
\begin{proof}
In the first case it is clear  that at all points $\pi_k$ which occurs
in between $\pi_i$ and $\pi_j$, $x$  has non-null demand as $x$ is used
directly in a strict computation.

In the second case, it can be seen that the liveness of $x$ due to the
\CONS\ is $\bcar\sigma$, where $\sigma$ is the demand on $z$. $\sigma$
definitely  includes  the  demand  $\acar\alpha$ due  to  its  use  in
$\pi_j$, therefore the demand on $x$ is at least $\alpha$.

 Since we do not kill the liveness of any variable along an evaluation
 path, and all program points along  an evaluation path share the same
 liveness environment, at any point along the evaluation path $x$ will
 not be set to $\bot$.
\end{proof}

For a program  executing according to {\em minefield  semantics}, if a
\bang\ has to occur then some variable which is dereferenced along the
evaluation path  must be set to  $\bot$, between its creation  and use
points.  This can  only happen  if the  demand on  the variable  being
dereferenced  was null.  From  the  lemma~\ref{lem:correctness} it  is
clear that this cannot happen.

The rest  of the proof depends  on converting a program  with function
calls to an equivalent program without function calls for which we 
refer the user to~\cite{asati14lgc}.

\begin{figure*}[h!]
\begin{center}
\begin{tabular}{|c|c|c|}
\hline
Premise & Transition & Rule name \\ 
\hline
\hline
          &\makecell{ $\rho, (\rho', x, e, \cred{\sigma', e_f'})\!:\!S,
  H, \kappa, \cred{\sigma, e_f}$ \\ $\rightsquigarrow \rho', S, H[\rho'(x) :=
    \kappa], e, \cred{\sigma', e_f'}$ }   &  \sc{const}
\\
\hline
          & \makecell{$\rho, (\rho', z, e, \cred{\sigma',
    e_f'})\!:\!S, H, (\CONS~x~y), \cred{\sigma, e_f}$ \\ $\rightsquigarrow
  \rho', S, H[\rho'(z) := (\rho(x),\rho(y))], e, \cred{\sigma', e_f'}$}     &  \sc{cons} \\
\hline
$H(\rho(x)) \mbox{ is } (l_1, l_2)$ & \makecell{$\rho, (\rho', z, e,
  \cred{\sigma', e_f'} )\!:\!S, H, (\CAR~x), \cred{\sigma, e_f}$ \\ $
  \rightsquigarrow \rho', S, H[\rho'(z) := H(l_1)], e, \cred{\sigma', e_f'}$}      &
\sc{car-whnf} \\
\hline
$H(\rho(x)) \mbox{ is } \langle s, \rho'\rangle$ &\makecell{ $\rho, S,
  H, (\CAR~x), \cred{\sigma, e_f}$ \\ $\rightsquigarrow \rho', (\rho, x,
  (\CAR~x), \cred{\sigma, e_f})\!:\!S, H, s, \cred{(\clazy \cup \acar)\sigma , \_ }$}      &
\sc{car-clo}
\\
\hline
$H(\rho(x)), H(\rho(y)) \in \mathbb{N}$
 & \makecell{$\rho, (\rho', z, e, \cred{\sigma', e_f'})\!:\!S, H,
  (+~x~y), \cred{\sigma, e_f}$ \\  $\rightsquigarrow \rho', S, H[\rho'(z)
    := H(\rho'(x)) + H(\rho'(y))], e, \cred{\sigma', e_f'}$}      &
\sc{prim-whnf} \\
\hline
$H(\rho(x)) \mbox{ is } \langle s, \rho'\rangle$ &\makecell{$\rho, S,
  H, (+~x~y), \cred{\sigma, e_f}$ \\ $\rightsquigarrow \rho', (\rho, x,
  (+~x~y), \cred{\sigma, e_f})\!:\!S, H, s, \cred{\clazy\sigma, \_}$}      &
\sc{prim-1-clo} \\
\hline
$H(\rho(y)) \mbox{ is } \langle s, \rho'\rangle $ & \makecell{$\rho,
  S, H, (+~x~y), \cred{\sigma, e_f}$ \\ $\rightsquigarrow \rho', (\rho, y,
  (+~x~y), \cred{\sigma, e_f})\!:\!S, H, s, \cred{\clazy\sigma, \_ }$}      &
\sc{prim-2-clo} \\
\hline
{$\mathit{g}~\mbox{defined as}$
$~(\DEFINE~(g~\myvec{y})~e_{\mathit{g}})$}  & \makecell{$\rho, S, H,
  (g~\myvec{x}), \cred{\sigma, e_f}$ \\ $\rightsquigarrow [\myvec{y} \mapsto
    \rho(\myvec{x})], S, H, e_{\mathit{g}}, \cred{\sigma, e_g}$}      &
\sc{funcall} \\
\hline
$\ell$ is a new location& \makecell{$\rho, S, H, (\LET~x\leftarrow
  s~\IN~e), \cred{\sigma, e_f}$ \\ $ \rightsquigarrow \rho\oplus[x \mapsto \ell], S, H[\ell \mapsto \langle s, \lfloor\rho\rfloor_{FV(s)}  \oplus [x \mapsto
  \ell]\rangle], e, \cred{\sigma, e_f}$} &
\sc{let} \\
\hline
$H(\rho(x)) \ne 0$ & \makecell{$\rho, S, H, (\pi:\SIF~\psi:x~e_1~e_2),
  \cred{\sigma, e_f}$ \\ $\rightsquigarrow \rho, S, H,  e_1, \cred{\sigma, e_f}$} & \sc{if-true} \\
\hline
$H(\rho(x)) = 0$ & \makecell{$\rho, S, H, (\pi:\SIF~\psi:x~e_1~e_2),
  \cred{\sigma, e_f}$ \\   $\rightsquigarrow
\rho, S, H,  e_2, \cred{\sigma, e_f}$} & \sc{if-false} \\
\hline
$H(\rho(x)) = \langle s, \rho' \rangle $ & \makecell{$\rho, S, H,
  (\pi:\SIF~\psi:x~e_1~e_2), \cred{\sigma, e_f}$ \\ $\rightsquigarrow
\rho', (\rho, x, (\SIF~x~e_1~e_2),  \cred{\sigma, e_f})\!:\!S, H, s,
\cred{\clazy\sigma, e_f}$}
&
\sc{if-clo} \\
\hline
{$H(\rho(x))~\mbox{is}$ $\mbox{whnf with value}~v$}& \makecell{$\rho,
  (\rho', z, e, \cred{\sigma', e_f'})\!:\!S, H, (\SRETURN~x), \cred{\sigma,
  e_f}$ \\ $\rightsquigarrow \rho', S, H[\rho'(z) := v], e,
  \cred{\sigma', e_f'}$} &
\sc{return-whnf}\\
\hline
$H(\rho(x)) = \langle s, \rho' \rangle $ & \makecell{$\rho, S, H,
  (\psi:\SRETURN~x), \cred{\sigma, e_f}$ \\ $
  \rightsquigarrow$
$\rho',~ (\rho, x, (\SRETURN~x), \cred{\sigma, e_f})\!:\!S, H,  s,
  \cred{\sigma, e_f}$} &
\sc{return-clo} \\
\hline
\end{tabular}
\caption{Minefield semantics for the language.\label{fig:minefield-semantics}}
\end{center}
\end{figure*}
\end{document}

>>>>

\documentclass[9pt]{sigplanconf}
\usepackage{etex}
\usepackage{float}
\usepackage{mathrsfs}
\usepackage{booktabs}
\usepackage{boxedminipage}
\usepackage[T1]{fontenc}
\usepackage{epsfig}
\usepackage{multirow}
\usepackage{url}
\usepackage[normalem]{ulem}
\usepackage{schemepgm}
\usepackage{graphicx}
\usepackage{graphpap}
\usepackage{tabularx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{pstricks}
\usepackage{pst-text}
\usepackage{pst-node}
\usepackage{pst-tree}
\usepackage{pst-rel-points}
\usepackage{bcprules}
\usepackage{subfigure}
\usepackage[boxed]{algorithm2e}
\usepackage{makecell}
\usepackage{balance}

@@

\subsection{Correctness of liveness analysis}

To set up the minefield semantics, we follow these steps:
\begin{enumerate}
\item We enrich  the abstract machine state $\rho,$ $S,$  $ H,$ $e$ to
  $\rho,$ $S,$ $H,$ $e,$ $\sigma,$ $e_f$, where $\sigma$ is the demand
  on the expression $e$, and $e_f$ is the body of the function. As the
  section on liveness  analysis shows, The entire  function body $e_f$
  and $\sigma$  are needed  to calculate  the liveness  environment at
  different program  points. The  demand sigma  is ``dynamic''  in the
  sense  that it  arises from  the actual  function calls  made during
  execution.   The  0-CFA  demand  is an  over-approximation  of  this
  demand.  The information  in a suspended evaluation  context is also
  similarly augmented with  the demand on it and the  function body of
  which the suspended evaluation is a  part.  Thus a stacked entry now
  takes the  form $(\rho, x, e,  \sigma, e_f)$. A modification  of the
  small step semantics which carries the extra information is shown in
  Figure~\ref{fig:minefield-semantics}.
\item Given  the current  evaluation context $\rho,  S, H,  e, \sigma,
  e_f$ and $LF$, we can construct a liveness environment as follows:
  \begin{enumerate}
  \item If $e$ is an expression, then:  
  $$ \Lv = \bigcup_{\psi \in \mathcal{P}(e)} \mathcal{M}(\psi),
    \text{where }
  \mathcal{M} = \mathcal{L}(e_f, \sigma, \Lfonly)$$
\item If $e$ is an application, then:  
  $$ \Lv = ref(e, \sigma, \Lfonly)$$
  \end{enumerate}
We can define in the same way the liveness environment for each of the
suspended  evaluation contexts in  $S$  giving  a  stack of  liveness
environments that we shall denote $\mathsf{SL}$.
\item $GC$ models a liveness based garbage collection:
$GC(\rho,$ $ S,$ $ H,$ $ L,$ $ SL) = (\rho', S', H')$ 
  \begin{enumerate}
  \item For each $x \in domain(\rho)$,  $\rho'(x)=\bot$ iff $x.\epsilon
    \notin \Lv$.
\item For each stack entry $(\rho,\_,\_,\_,\_)$ in $S$ with $\Lv'$ as
  the corresponding liveness environment in $\mathsf{SL}$,  and for each $x
  \in domain(\rho)$,  $\rho'(x)=\bot$
  iff $x.\epsilon \notin \Lv'$.
\item For each location $\ell$, $H'(\ell) = \bot$, iff  there is no
  $x$ in either the current environment or one of the stacked
  environments $\rho$ such that for some $\alpha$, $H[x.\alpha] =
  \ell$  and $x.\alpha \in \Lv$, .
  \end{enumerate}
\end{enumerate}

Starting   from  the   initial  state   $([\;]_\rho,([\;]_\rho,  \ans,
(\print~\ans)):[\;]_{S} ,  [\;]_{H}, (\mainpgm)$, every  transition is
preceded by  a garbage collection  using $GC$.  We consider  a garbage
collection followed  by a transition as  a step. We show  by induction
that, for programs  without function calls, starting  from the initial
state that  has $e_{\main}$  as the first  expression to  be evaluated
instead  of $\main$,  any transition  of  $n$ steps  occurs without  a
\bang.

Observe  that  the  transitions  that dereference  and  therefore  can
potentially lead to \bang\ are those  which have a $H(\ldots)$ term in
the  transition  rule.   Of  these,  terms  of  the  form  $H(\ldots)$
dereference a root  variable, and the term $H(l)$  dereferences a heap
location.  In  both the  cases we show  that the  dereferencing cannot
lead to a \bang\ state in programs without function calls.

We  need the  following proposition that  connects  liveness analysis  with
minefield semantics transitions. 
\begin{proposition}
\label{prop:correctness}
\begin{enumerate} 
\item Consider a \LET\ definition $:\pi: (\LET~x
  \leftarrow~\ldots~\IN~e$, such that there is a use of $x$
  in $e$, i.e.  an occurrence of $\LET~y~\leftarrow~s~in
  \ldots$, where $s$ is $(\CAR\ x)$, $(\CDR\ x)$ or $(+\ x\ y)$, or
  $(\SIF\ x\ e1\ e2)$ or $(\SRETURN\ x)$. Then the demand on $x$ at $\pi$ is non-null. 
\item Assume that a  \LET\ definition $\pi: (\LET~z  \leftarrow~(\CONS~x~y)~\IN~e$ has an occurrence of 
 $(\CAR~z)$ in $e$. Then the liveness of $x$ at $\pi$ includes $\epsilon$.  Similarly if $e$ contains $(CDR~z)$, then the liveness of$y$ at $\pi$ includes $\epsilon$.  
\end{enumerate},
\end{proposition}

In  the second  case for  example,  the liveness  of $x$  due to  the
\CONS\ is  $\bcar\sigma$, where $\sigma$  is the demand on  $z$. Since
$\sigma$  includes  the  demand  $\acar\alpha$   due  to  its  use  in
$(\CAR~z)$, the liveness on $x$  includes $\epsilon$. We thus have the
following lemma:

\begin{lemma}\label{lemma:no-calls-correctness} A program without function calls cannot enter the \bang\ state in the minefield semantics.
\end{lemma}

\begin{proof}
For a program  executing according to {\em minefield  semantics}, if a
\bang\ has to  occur then some variable which is  dereferenced must be
set to  $\bot$, between its  creation and  use points.  This  can only
happen if  the liveness of  the variable being dereferenced  was null.
From  the lemma~\ref{lem:correctness}  it  is clear  that this  cannot
happen.
\end{proof}

For   a    programs   with   functions,   we    use   the   techniques
in~\cite{asati14lgc}. We  assume as  a contradiction that  the program
can enter a \bang\ state  and inline the  functions called to  obtain a
program that is without functions  but has the same execution behavior
and    demands    as   the    original    program.     We   then    use
Lemma~\ref{lemma:no-calls-correctness}  to  show  that  the  resulting
functionless program cannot enter a bang state.

@@

\balance
\bibliography{fun_hra}{}
\bibliographystyle{abbrv}

\appendix

\section{Minefield semantics for the language}

\begin{figure*}[h!]
\begin{center}
\begin{tabular}{|c|c|c|}
\hline
Premise & Transition & Rule name \\ 
\hline
\hline
          &\makecell{ $\rho, (\rho', x, e, \cred{\sigma', e_f'})\!:\!S,
  H, \kappa, \cred{\sigma, e_f}$ \\ $\rightsquigarrow \rho', S, H[\rho'(x) :=
    \kappa], e, \cred{\sigma', e_f'}$ }   &  \sc{const}
\\
\hline
          & \makecell{$\rho, (\rho', z, e, \cred{\sigma',
    e_f'})\!:\!S, H, (\CONS~x~y), \cred{\sigma, e_f}$ \\ $\rightsquigarrow
  \rho', S, H[\rho'(z) := (\rho(x),\rho(y))], e, \cred{\sigma', e_f'}$}     &  \sc{cons} \\
\hline
$H(\rho(x)) \mbox{ is } (l_1, l_2)$ & \makecell{$\rho, (\rho', z, e,
  \cred{\sigma', e_f'} )\!:\!S, H, (\CAR~x), \cred{\sigma, e_f}$ \\ $
  \rightsquigarrow \rho', S, H[\rho'(z) := H(l_1)], e, \cred{\sigma', e_f'}$}      &
\sc{car-whnf} \\
\hline
$H(\rho(x)) \mbox{ is } \langle s, \rho'\rangle$ &\makecell{ $\rho, S,
  H, (\CAR~x), \cred{\sigma, e_f}$ \\ $\rightsquigarrow \rho', (\rho, x,
  (\CAR~x), \cred{\sigma, e_f})\!:\!S, H, s, \cred{(\clazy \cup \acar)\sigma , \_ }$}      &
\sc{car-clo}
\\
\hline
$H(\rho(x)), H(\rho(y)) \in \mathbb{N}$
 & \makecell{$\rho, (\rho', z, e, \cred{\sigma', e_f'})\!:\!S, H,
  (+~x~y), \cred{\sigma, e_f}$ \\  $\rightsquigarrow \rho', S, H[\rho'(z)
    := H(\rho'(x)) + H(\rho'(y))], e, \cred{\sigma', e_f'}$}      &
\sc{prim-whnf} \\
\hline
$H(\rho(x)) \mbox{ is } \langle s, \rho'\rangle$ &\makecell{$\rho, S,
  H, (+~x~y), \cred{\sigma, e_f}$ \\ $\rightsquigarrow \rho', (\rho, x,
  (+~x~y), \cred{\sigma, e_f})\!:\!S, H, s, \cred{\clazy\sigma, \_}$}      &
\sc{prim-1-clo} \\
\hline
$H(\rho(y)) \mbox{ is } \langle s, \rho'\rangle $ & \makecell{$\rho,
  S, H, (+~x~y), \cred{\sigma, e_f}$ \\ $\rightsquigarrow \rho', (\rho, y,
  (+~x~y), \cred{\sigma, e_f})\!:\!S, H, s, \cred{\clazy\sigma, \_ }$}      &
\sc{prim-2-clo} \\
\hline
{$\mathit{g}~\mbox{defined as}$
$~(\DEFINE~(g~\myvec{y})~e_{\mathit{g}})$}  & \makecell{$\rho, S, H,
  (g~\myvec{x}), \cred{\sigma, e_f}$ \\ $\rightsquigarrow [\myvec{y} \mapsto
    \rho(\myvec{x})], S, H, e_{\mathit{g}}, \cred{\sigma, e_g}$}      &
\sc{funcall} \\
\hline
$\ell$ is a new location& \makecell{$\rho, S, H, (\LET~x\leftarrow
  s~\IN~e), \cred{\sigma, e_f}$ \\ $ \rightsquigarrow \rho\oplus[x \mapsto \ell], S, H[\ell \mapsto \langle s, \lfloor\rho\rfloor_{FV(s)}  \oplus [x \mapsto
  \ell]\rangle], e, \cred{\sigma, e_f}$} &
\sc{let} \\
\hline
$H(\rho(x)) \ne 0$ & \makecell{$\rho, S, H, (\pi:\SIF~\psi:x~e_1~e_2),
  \cred{\sigma, e_f}$ \\ $\rightsquigarrow \rho, S, H,  e_1, \cred{\sigma, e_f}$} & \sc{if-true} \\
\hline
$H(\rho(x)) = 0$ & \makecell{$\rho, S, H, (\pi:\SIF~\psi:x~e_1~e_2),
  \cred{\sigma, e_f}$ \\   $\rightsquigarrow
\rho, S, H,  e_2, \cred{\sigma, e_f}$} & \sc{if-false} \\
\hline
$H(\rho(x)) = \langle s, \rho' \rangle $ & \makecell{$\rho, S, H,
  (\pi:\SIF~\psi:x~e_1~e_2), \cred{\sigma, e_f}$ \\ $\rightsquigarrow
\rho', (\rho, x, (\SIF~x~e_1~e_2),  \cred{\sigma, e_f})\!:\!S, H, s,
\cred{\clazy\sigma, e_f}$}
&
\sc{if-clo} \\
\hline
{$H(\rho(x))~\mbox{is}$ $\mbox{whnf with value}~v$}& \makecell{$\rho,
  (\rho', z, e, \cred{\sigma', e_f'})\!:\!S, H, (\SRETURN~x), \cred{\sigma,
  e_f}$ \\ $\rightsquigarrow \rho', S, H[\rho'(z) := v], e,
  \cred{\sigma', e_f'}$} &
\sc{return-whnf}\\
\hline
$H(\rho(x)) = \langle s, \rho' \rangle $ & \makecell{$\rho, S, H,
  (\psi:\SRETURN~x), \cred{\sigma, e_f}$ \\ $
  \rightsquigarrow$
$\rho',~ (\rho, x, (\SRETURN~x), \cred{\sigma, e_f})\!:\!S, H,  s,
  \cred{\sigma, e_f}$} &
\sc{return-clo} \\
\hline
\end{tabular}
\caption{Minefield semantics.\label{fig:minefield-semantics}}
\end{center}
\end{figure*}

\end{document}
